### 2016-03-23 Draft.tex
Implementing some of Jarad's comments.

\documentclass[useAMS,usenatbib,referee,12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage[round]{natbib}
\usepackage{color}
\usepackage{caption}
\usepackage{tabularx, booktabs}
\usepackage{float}
\usepackage{adjustbox}
\usepackage[toc,page]{appendix}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{setspace}
\doublespacing

\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
\newcommand{\jarad}[1]{{\color{red} #1}}
\newenvironment{indpar}[1]%
     {\begin{list}{}%
             {\setlength{\leftmargin}{#1}}%
             \item[]%
     }
     {\end{list}}

@@

Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate.  However, the assumption of constant detection rates during surveys is often not justified.  Detection rates can be influenced by organismal behaviors, including responses to observer presence and heterogeneity among population subgroups, and also by survey methods that affect observer effort.  Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance. \jarad{in the abstract focus on what we did} Under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals.  We instead model the detection {\it rate} in continuous time via a time-to-detection distribution embedded within a hierarchical N-mixture framework using Bayesian methods.  Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  We apply this model to Ovenbird counts from the Chippewa National Forest and to datasets simulated under different time-to-detection patterns.  Models assuming constant detection rates produce biased estimates of detection when true detection rates vary with time, whereas models allowing for variable detection (assuming gamma, Weibull, or lognormal distributed times to detection) produce less biased estimates of the detection probability and nominal credible interval coverage.  Models ignoring detection heterogeneity across subgroups yield biased estimates of detection when such heterogeneity exists, whereas models accounting for detection heterogeneity (modeled as a mixture) return reasonable coverage rates and can outperform heterogeneity-ignorant models even when there is no heterogeneity.

@@

Increasing hazard functions result in a probability distribution function (pdf) for $T$ that has a mode away from zero; we term these TTDDs `peaked'.  We contrast these to decreasing hazard functions with a mode at zero, which we term `non-peaked'.  For our purposes, the exponential TTDD is neither peaked nor non-peaked.

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.68\textwidth]{Hazards}
\caption{Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs. \jarad{Are these setup to have the same mean and variance or anything like that?}}
\label{HazardFxns}
\end{figure}

@@

Our model is formulated for single-visit removal-sampled point-count surveys conducted at a set of sites $s=1,\dotso,S$ repeated across distinct time periods $r=1,\dotso,R$. \jarad{So time periods are R because we already use t for intervals? Argh!}  We present the model for times to detection censored into $i=1,\dotso,I$ intervals \jarad{Why don't we present the exact time-to-detection model first and then censoring it?}, but it is straight-forward to simplify the model for exact time-to-detection data.  
We denote the number of individual observed at site $s$, time $r$, and in interval $i$ as $n_{sri}$ and define the total number of individuals observed during survey as $\vn_{sr} = (\sum_{i=1}^I n_{sri})n_{sr1},\ldots,n_{srI})$.

@@

We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a multinomial data model, (b) a detection model based upon the partitioned cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.  Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.

\jarad{We should try to avoid using subsubsections. Perhaps we should provide a data section and then a model section?}

\subsubsection{Data Model}
We assume that counts from different surveys are conditionally independent.  We also assume that the time-to-detection is independent, identically distributed for all individuals of a species at any given survey.  Given that an organism is detected, we denote the probability it is first detected during interval $i$ as $p_{sri}$ such that $\sum_{i=1}^I p_{sri} = 1$.  We write $\textbf{p}_{sr}$ to denote the vector of interval-specific detection probabilities at any given survey.  The data model is then:
\begin{align}\label{eq:datamodel}
\vn_{sr} &\sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)\\
n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)
\end{align}

\jarad{Pedogogically, I would introduce the binomial model first and then the multinomial model. Or we could introduce a Multinomial with parameter $N_{sr}$ and an additional bin for non-detection.}

where for each survey, $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$ is the number of individuals counted across all intervals, $N_{sr}$ is the total number of individuals present (counted and uncounted), and $p_{sr}^{(det)}$ is the probability that an individual is detected at any time during the observation period.  The number of uncounted individuals at any survey is then: $n_{sr}^{(unobs)} = N_{sr} - n_{sr}^{(obs)}$.

\subsubsection{Detection Model}\label{sec:detectionmodel}

@@

For interval-censored data, we denote the cutoffs for each observation interval with $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  Conditioned on an individual being detected \jarad{Would this be easier to introduce without conditioning on individuals being detected?}, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.  In summary, the detection model is:
\begin{align}
p_{sri} &= \left[F_T(C_i|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) - F_T(C_{i-1}|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\right]\big/p_{sr}^{(det)}\\
\label{eq:pdet}p_{sr}^{(det)} &= F_T(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\\
F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) &= \text{a TTDD as described above}
\end{align}

@@

\subsubsection{Abundance Model}

@@

\subsubsection{Priors}\label{sec:priors}
Our goal in selecting priors is to be non-informative within a reasonable range of values.  Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.  Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.  For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.  All priors are independent.
\jarad{This takes up too much space.}
\begin{align*}
\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
\text{Intercept}^A &\sim N(1, 1)\\
\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
\boldsymbol{\beta}^A &\sim N(0, 1)\\
\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
\gamma &\sim \text{Beta}(1,1)
\end{align*}

@@

\subsubsection{Mixture vs. non-mixture TTDD}\label{sec:mixture}

@@

\subsubsection{Constant vs. non-constant detection}\label{sec:family}

@@

\subsubsection{Simulations with covariates and random effects}\label{sec:simfull}

@@

\subsubsection{Model fit}\label{sec:modelfit}

>>>>

\documentclass[useAMS,usenatbib,referee,12pt]{article}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{parskip}
\usepackage[round]{natbib}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage{caption}
\usepackage{tabularx, booktabs}
\usepackage{float}
\usepackage{adjustbox}
\usepackage[toc,page]{appendix}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{setspace}
\doublespacing

\newcommand{\adam}[1]{{\color{blue} ADAM: #1}}
\newcommand{\jarad}[1]{{\color{Orange} #1}}
\newenvironment{indpar}[1]%
     {\begin{list}{}%
             {\setlength{\leftmargin}{#1}}%
             \item[]%
     }
     {\end{list}}

@@

Abundance estimates from animal point-count surveys require accurate estimates of detection probabilities.  The standard model for estimating detection from removal-sampled point-count surveys assumes that organisms at a survey site are detected at a constant rate; however, this assumption is often not justified.  Detection rates can be influenced by organismal behaviors, survey methodology, and observer effort.  Failure to account for non-constant detection leads to biased estimates of detection and therefore of abundance.  Whereas, under the standard approach, detection probability is modeled by dividing the observation period into equal-duration intervals, we instead model the detection {\it rate} in continuous time via a time-to-detection distribution embedded within a hierarchical N-mixture framework using Bayesian methods.  Our model is thus a combination of survival time-to-event analysis with unknown-N, unknown-p abundance estimation.  We apply this model to Ovenbird counts from the Chippewa National Forest and to datasets simulated under different time-to-detection patterns.  Models assuming constant detection rates produce biased estimates of detection when true detection rates vary with time, whereas models allowing for variable detection (assuming gamma, Weibull, or lognormal distributed times to detection) produce less biased estimates of the detection probability and nominal credible interval coverage.  Models ignoring detection heterogeneity across subgroups yield biased estimates of detection when such heterogeneity exists, whereas models accounting for detection heterogeneity (modeled as a mixture) return reasonable coverage rates and can outperform heterogeneity-ignorant models even when there is no heterogeneity.

@@

Increasing hazard functions result in a pdf for $T$ that has a mode away from zero; we term these TTDDs `peaked'.  We contrast these to decreasing hazard functions with a mode at zero, which we term `non-peaked'.  For our purposes, the exponential TTDD is neither peaked nor non-peaked.

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.68\textwidth]{Hazards}
\caption{Examples of instantaneous rates of detection (hazard functions) from exponential, gamma, lognormal, and Weibull TTDDs.}
\label{HazardFxns}
\end{figure}

@@

Our model is formulated for single-visit removal-sampled point-count surveys at a set of sites $s=1,\dotso,S$ repeated across distinct sampling periods $r=1,\dotso,R$.  We present the model for exact time-to-detection data and also show how to modify it for interval-censored detection times.  For exact data, we use $n_{sr}^{(obs)}$ to denote the number of individuals observed during a survey, each with a time to detection $t_{srq}$, where $q = 1,\dotso,n_{sr}^{(obs)}$ indexes observed individuals.  For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  The number of animals counted during any interval is $n_{sri}$, the vector of counts for a survey is written $\vn_{sr}$, and $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$.

@@

We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a binomial data model, (b) a detection model based upon the cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.  Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.

\subsection{Data Model}
We assume that counts from different surveys are conditionally independent.  We also assume that the times to detection for all individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$.  The probability of detecting an animal is thus $p_{sr}^{(det)} = Pr(t_{srq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{srq}<C_I) = f_T(t) / F_T(C_I)$.  If the total number of animals present at the survey is $N_{sr}$, then the data model is:
\begin{align}\label{eq:datamodel}
n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)\\
\label{eq:Tsrq}T_{srq} &\sim f_{T|det}(t|t_{srq}<C_I)
\end{align}
When detection times are censored, we replace Eq. \ref{eq:Tsrq} with $\vn_{sr} \sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)$, where $\textbf{p}_{sr}$ is the vector of interval-specific detection probabilities obtained by partitioning the conditional cdf $F_{T|det}(t|t_{srq}<C_I)$ at the interval cutoff times.  It follows that $\sum_{i=1}^I p_{sri} = 1$.

\subsection{Detection Model}\label{sec:detectionmodel}

@@

Conditioned on an individual being detected \jarad{Would this be easier to introduce without conditioning on individuals being detected?}, we calculate interval-specific detection probabilities $p_{sri}$ by partitioning the cdf of the TTDD at the observation interval cutoffs and normalizing by the binomial detection probability $p_{sr}^{(det)}$.  In summary, the detection model is:
\begin{align}
p_{sri} &= \left[F_T(C_i|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) - F_T(C_{i-1}|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\right]\big/p_{sr}^{(det)}\\
\label{eq:pdet}p_{sr}^{(det)} &= F_T(C_I|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta})\\
F_T(t|\textbf{X}_{sr}^D, \textbf{Z}_{sr}^D, \boldsymbol{\theta}) &= \text{a TTDD as described above}
\end{align}

@@

\subsection{Abundance Model}

@@

\subsection{Priors}\label{sec:priors}
Our goal in selecting priors is to be non-informative within a reasonable range of values.  Priors for Intercept$^D$ are centered so that the median detection probability is $p_{sr}^{(det)} = 0.50$, based on an intercept-only non-mixture model with scale parameter 1.  Standard deviations are chosen so that 95\% of the prior detection probability rests within $p_{sr}^{(det)} \in (0.01, 1.0)$.  Priors for Intercept$^A$ center at a median abundance of about 3 birds per site and a 95\% probability of 0-14 birds present (counted and uncounted).  Because covariates are centered and scaled, priors for effect parameters are centered at zero with standard deviations matching the appropriate intercept term.  For all shape and variance parameters, we choose half-Cauchy(0,1) priors, and for the mixing parameter $\gamma$ we chose a uniform prior.  All priors are independent.
\jarad{This takes up too much space.}
\begin{align*}
\text{Intercept}^D &\sim N(-2.7, 2.25^2) \text{ or } N(-2.3, 1.2^2)\\
\text{Intercept}^A &\sim N(1, 1)\\
\boldsymbol{\beta}^D &\sim N(0, 2.25^2)\\
\boldsymbol{\beta}^A &\sim N(0, 1)\\
\sigma_D, \sigma_A, \alpha, \sigma_{det}, k &\sim \text{half-Cauchy}(0,1)\\
\gamma &\sim \text{Beta}(1,1)
\end{align*}

@@

\subsection{Mixture vs. non-mixture TTDD}\label{sec:mixture}

@@

\subsection{Constant vs. non-constant detection}\label{sec:family}

@@

\subsection{Simulations with covariates and random effects}\label{sec:simfull}

@@

\subsection{Model fit}\label{sec:modelfit}
